First Exclusive Measurement of Deep Virtual Compton Scattering off $^4$He: Toward the 3D tomography of nuclei

### 2017-02-13 eg6-letters/coherent/dvcs_he4_prl.tex
update the paper with the remaining suggestions from Stepan

%asymmetries
In this work, the physics observable extracted using coherent DVCS events is
the Beam Spin Asymmetry (BSA). The BSA on unpolarized target (U), $A_{LU}$, is 
measurable by the difference of cross sections of the reaction at opposite beam 
helicities normalized to the total cross section. It is written as:
  \begin{equation}
  A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
                {d^{5}\sigma^{+} + d^{5}\sigma^{-}}.
    \label{BSA_equation}
  \end{equation}
where $d^{5}\sigma^{+(-)}$ is the DVCS differential cross 
section for a positive (negative) beam helicity. Experimentally, $A_{LU}$ 
can be expressed in terms of the measured number of events in each 
beam-helicity state ($N^{+}$, $N^{-}$) as:
\begin{equation}
A_{LU} = \frac{1}{P_{B}} \frac{N^{+} - N^{-}}{N^{+} + N^{-} }.
\end{equation}
where $P_{B}$ is the beam polarization, and $N^{+}$ and $N^{-}$ are the number 
of DVCS events detected with positive and negative electron helicity with 
respect to the beam direction. In this ratio most of uncertainties of the 
experimental normalization and in detector efficiencies will cancel out.

To interpret DVCS, one has to keep in mind that this reaction is 
indistinguishable from the Bethe-Heitler (BH) process, where the final photon 
is emitted either from the incoming or the outgoing leptons. At leading twist 
and leading order, the deeply virtual photon production cross section is 
composed of BH, DVCS, and interference terms. The amplitudes of the 
three terms can be approximated as a finite sum of Fourier harmonics, as shown 
for the nucleon in \cite{Belitsky:2001ns} and for the spin-zero targets in 
\cite{Kirchner:2003wt,Belitsky:2008bz}. Based on these calculations, the 
beam-spin asymmetry ($A_{LU}$) of a spin-zero hadron can then be expressed as:
\small
\begin{equation}
\begin{split}
A_{LU}(\phi) =~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\\
 \frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( 
\Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)}
\label{eq:A_LU-coh}
\end{split}
\end{equation}
\normalsize
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the $^4$He CFF $\mathcal{H}_{A}$ associated with the GPD $H_A$,
and $\phi$ is the azimuthal angle between the ($e$,$e'$) and 
($\gamma^{*}$,$^4$He$'$) planes. The $\alpha_{i}$'s are $\phi$-dependent 
kinematical factors that depend on the nuclear form factor ($F_{A}(t)$) and the 
independent variables $Q^2$, $x_{B}$ and $t$. These factors are expressed as:
\small
\begin{eqnarray}
   \alpha_0 (\phi) & = &\frac{x_{A}(1+\epsilon^2)^2}{y} S_{++}(1) \sin(\phi)\\
    \alpha_1 (\phi) & = & c_0^{BH}+c_1^{BH} \cos({\phi})+c_2^{BH} \cos(2\phi)\\ 
   \alpha_2 (\phi) & = & \frac{x_{A}(1+\epsilon^2)^2}{y}  \left( C_{++}(0) +  
C_{++}(1) cos(\phi) \right)\\
\alpha_3 (\phi) &=& \frac{x^{2}_{A}t(1+\epsilon^2)^2}{y} {\mathcal P}_1(\phi) 
{\mathcal P}_2(\phi) \cdot 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + 
\epsilon^2}
\end{eqnarray}
\normalsize
Where $x_{A} = \frac{x_{B}M_{N}}{M_{A}}$ with $M_{A}$ the $^4$He mass and 
$\mathcal{P}_1(\phi)$ and $\mathcal {P}_2(\phi)$ the Bethe-Heitler
propagators. The factors $c_{0,1,2}^{BH}$, $c_0^{DVCS}$, $c_{0,1}^{INT}$ and
$s_1^{INT}$ are the Fourier coefficients of the BH, and $S{++}(1)$, $C_{++}(0)$, 
and $C_{++}(1)$ are the Fourier harmonics in the leptonic tensor. The 
explicit expressions of these terms can be found in \cite{Belitsky:2008bz}. 
We see that by using the $\sin(\phi)$ and $\cos(\phi)$ contributions, it is 
possible to extract 
$\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ from the beam spin
asymmetry.

Due to limited statistics, we only bin  our data two-dimensionally 
 in the azimuthal angle ($\phi$) and one of the kinemtical variables 
$Q^{2}$, $x_{B}$ and $t$. Figure \ref{fig:alu} presents $A_{LU}$ for the three
sets of binning. The asymmetries are fitted with the form of
equation \ref{eq:A_LU-coh}, where the real and imaginary part of the CFF
$\mathcal{H}_{A}$ are the free parameters of the fit. We identified two main 
backgrounds to our measurement, accidental coincidences and exclusive, coherent 
$\pi^0$ production. The accidentals have particles originating from different 
events, and we estimate their  contribution to be 4.1\% of our sample. We 
evaluated this contribution by selecting events passing all our cuts but with 
an electron and helium originating from different vertices. Regarding the 
$\pi^0$ production, it can easily be mistaken for DVCS when one of the two 
photons from the $\pi^0$ decay is produced at low energy in the laboratory 
frame and remains undetected. To estimate the importance of this background, we 
developed an event generator calibrated on the experimental yield of exclusive 
$\pi^0$ measured by our experiment. We used this generator together with a 
GEANT3 simulation of our detector to estimate the ratio between the number of 
$\pi^0$ events where the two photons are detected and those that is 
misidentified as DVCS events. This ratio is then multiplied by the measured 
yield of exclusive $\pi^0$ events to correct the DVCS data. Depending on the 
kinematics, we found contaminations of 2 to 4\%. The study of systematics 
uncertainties showed that the main contributions come from the choice of the 
DVCS exclusivity cuts (8\%) and the large binning size (5.1\%). However, added 
quadratically,the total systematic uncertainty is about 10\%, which is 
significantly smaller than statistical uncertainties in all kinematical bins. 
Figure \ref{fig:alu90} shows the $Q^2$, $x_{B}$, and $-t$-dependencies of the 
fitted $A_{LU}$ signals at $\phi$~=~90$^{\circ}$. The $x_{B}$ and 
$-t$-dependencies are compared to theoretical calculations performed by 
S.~Liuti and K.~Taneja \cite{simonetta_2}. Their model relies on the impulse 
approximation and uses advanced spectral functions of nuclei. The calculations 
are at slightly different kinematics than our data but still provide some 
guidance. The experimental results appear to have larger asymmetries than the 
calculations. These differences may arise from nuclear effects which are not 
taken into account in the model, such as long-range interactions.

@@

We display in figure \ref{fig:CFF_HA} our extraction of the $^4$He CFF 
$\mathcal{H}_A$, together with theoretical calculations from two models, 
labelled {\em convolution} and {\em off-shell}. In the convolution 
model \cite{Vadim_priv}, the nucleus is assumed to be composed of 
non-relativistic and non-interacting nucleons, where these nucleons interact 
independently with the probe. The Convolution-Dual model is based on nucleon 
GPDs from the dual parametrization \cite{Guzey:2006xi}, where the 
Convolution-VGG uses nucleon GPDs  from the VGG model and is based on the 
double distributions ansatz \cite{DD_model}. The off-shell model 
\cite{GonzalezHernandez:2012jv} relies on the impulse approximation and uses 
advanced spectral functions of  nuclei that account for all configurations of 
the final nuclear system and the binding effects between the nucleons.

The results in Figure \ref{fig:CFF_HA} show that the extraction of the CFF
from the beam spin asymmetry is possible without any model dependent 
assumptions. The amplitude and the dependencies observed as a function of 
$Q^{2}$, $x_B$, and $-t$ are in agreement 
with the theoretical calculations. One can see a difference between the 
precision of the extracted imaginary and real parts, which is expected from
equation \ref{eq:A_LU-coh} because $\alpha_2$ is smaller than $\alpha_0$. 
While these results are not precise enough to discriminate between the models,
they are very promising for future experiments and demonstrate our ability
to directly extract the CFF of the helium nucleus in a completely model
independent way.

%conclusion
In summary, we present the first exclusive measurement of coherent DVCS 
off $^4$He using the CLAS spectrometer supplemented  
by a specially designed RTPC. This dataset represents a 
unique source as the first exclusive coherent nuclear DVCS measurement which 
can be used to directly constrain GPD models. The measured beam-spin asymmetries 
show a signal stronger than for the proton and allow to perform a fully 
model-independent extraction of the $^4$He CFF $\mathcal{H}_A$. The extracted 
CFF is in a good agreement with the available models. This opens 
many new perspectives to study the nuclear structure within the GPD framework 
and paves the way for future, more precise measurements at JLab 12~GeV to 
eventualy localize nuclear effects in the transverse plane.

>>>>

%asymmetries
In this work, the physics observable extracted using coherent DVCS events is
the Beam Spin Asymmetry (BSA). The BSA on unpolarized target (U), $A_{LU}$, is 
measurable by the difference of cross sections of the reaction at opposite beam 
helicities normalized to the total cross section. It is written as:
  \begin{equation}
  A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
                {d^{5}\sigma^{+} + d^{5}\sigma^{-}},
    \label{BSA_equation}
  \end{equation}
where $d^{5}\sigma^{+(-)}$ is the DVCS differential cross 
section for a positive (negative) beam helicity. Experimentally, $A_{LU}$ 
can be expressed in terms of the measured number of events in each 
beam-helicity state ($N^{+}$, $N^{-}$) as:
\begin{equation}
A_{LU} = \frac{1}{P_{B}} \frac{N^{+} - N^{-}}{N^{+} + N^{-} },
\end{equation}
where $P_{B}$ is the beam polarization, and $N^{+}$ and $N^{-}$ are the number 
of DVCS events detected with positive and negative electron helicity with 
respect to the beam direction. In this ratio most of uncertainties of the 
experimental normalization and in detector efficiencies will cancel out.

The DVCS reaction is experimentally indistinguishable from the well-known 
Bethe-Heitler (BH) process, where the final photon is emitted either from the 
incoming or the outgoing leptons. At leading twist and leading order, the 
deeply virtual photon production cross section is composed of BH, DVCS, and 
interference terms. While the BH is not sensitive to GPDs and does not carry 
information about the partonic structure of the hadronic target, its 
interference with the DVCS reaction is sensitive to GPDs enhancing the DVCS 
small cross section at the kinematical regions explored at Jefferson Lab. The 
amplitudes of the three terms can be approximated as a finite sum of Fourier 
harmonics, as shown for the nucleon in \cite{Belitsky:2001ns} and for the 
spin-zero targets in \cite{Kirchner:2003wt,Belitsky:2008bz}. Based on these 
approximations, the beam-spin asymmetry ($A_{LU}$) of a spin-zero hadron can 
then be expressed as:
\small
\begin{equation}
\begin{split}
A_{LU}(\phi) =~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\\
 \frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( \Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)},
\label{eq:A_LU-coh}
\end{split}
\end{equation}
\normalsize
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the $^4$He CFF $\mathcal{H}_{A}$ that depends on the GPD 
$H_A$. In Eq.~\ref{eq:A_LU-coh}, $\phi$ is the azimuthal angle between the 
($e$,$e'$) and ($\gamma^{*}$,$^4$He$'$) planes. The $\alpha_i$’s are kinematic 
factors that depend on azimuthal angle $\phi$, the nuclear form factor 
$F_A(t)$, $Q^2$, and $x_B$. These factors are expressed as:
\small
\begin{eqnarray}
   \alpha_0 (\phi) & = &\frac{x_{A}(1+\epsilon^2)^2}{y} S_{++}(1) \sin(\phi),\\
    \alpha_1 (\phi) & = & c_0^{BH}+c_1^{BH} \cos({\phi})+c_2^{BH} 
   \cos(2\phi),\\ \alpha_2 (\phi) & = & \frac{x_{A}(1+\epsilon^2)^2}{y}  \left( 
C_{++}(0) +  C_{++}(1) cos(\phi) \right),\\
\alpha_3 (\phi) &=& \frac{x^{2}_{A}t(1+\epsilon^2)^2}{y} {\mathcal P}_1(\phi) 
{\mathcal P}_2(\phi) \cdot 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + 
\epsilon^2},
\end{eqnarray}
\normalsize
where $x_{A} = \frac{x_{B}M_{N}}{M_{A}}$ with $M_{A}$($M_{N}$) is the $^4$He 
(nucleon) mass, $\epsilon = \frac{2x_{A}M_{A}}{\sqrt{Q^{2}}}$ and 
$y=\frac{Q^{2}}{2x_{A}M_{A}E_{beam}}$, $\mathcal{P}_1(\phi)$ and $\mathcal 
{P}_2(\phi)$ are the Bethe-Heitler propagators. The factors $c_{0,1,2}^{BH}$, 
$c_0^{DVCS}$, $c_{0,1}^{INT}$ and $s_1^{INT}$ are the Fourier coefficients of 
the BH, and $S_{++}(1)$, $C_{++}(0)$, and $C_{++}(1)$ are the Fourier harmonics 
in the leptonic tensor.  The explicit expressions of these terms can be found 
in \cite{Belitsky:2008bz}.  We see that by using the $\sin(\phi)$ and 
$\cos(\phi)$ contributions, it is possible to extract $\Im m(\mathcal{H}_{A})$ 
and $\Re e(\mathcal{H}_{A})$ from the beam spin
asymmetry.

In this work the azimuthal dependence of the BSA, $A_{LU}$, has been studied 
for a wide range of kinematics. We identified two main backgrounds to our 
measurement, accidental coincidences and exclusive coherent $\pi^0$ production.  
The accidentals have particles originating from different events, and we 
estimate their  contribution to be 4.1\% of our sample. We evaluated this 
contribution by selecting events passing all our cuts but with an electron and 
helium originating from different vertices. Regarding the $\pi^0$ production, 
it can easily be mistaken for DVCS when one of the two photons from the $\pi^0$ 
decay is produced at low energy in the laboratory frame and remains undetected.  
To estimate the importance of this background, we developed an event generator 
calibrated on the experimental yield of exclusive $\pi^0$ measured by our 
experiment. We used this generator together with a GEANT3 simulation of our 
detector to estimate the ratio between the number of $\pi^0$ events where the 
two photons are detected and those that is misidentified as DVCS events. This 
ratio is then multiplied by the measured yield of exclusive $\pi^0$ events to 
correct the DVCS data.  Depending on the kinematics, we found contaminations of 
2 to 4\%. The study of systematics uncertainties showed that the main 
contributions come from the choice of the DVCS exclusivity cuts (8\%) and the 
large binning size (5.1\%). However, added quadratically,the total systematic 
uncertainty is about 10\%, which is significantly smaller than statistical 
uncertainties in all kinematical bins.

Due to limited statistics, dependence on the kinematical variables $Q^2$, 
$x_B$, and $t$ has been studied separately. In Figure \ref{fig:alu}, $A_{LU}$ 
for the three sets of binning is presented. The curves on the plots are fits 
with the function presented in Eq.~\ref{eq:A_LU-coh}, where the real and 
imaginary part of the CFF $\mathcal{H}_{A}$ are the free parameters of the fit. 
In Figure \ref{fig:alu90} the $Q^2$, $x_{B}$, and $-t$-dependencies of the 
fitted $A_{LU}$ signals at $\phi$~=~90$^{\circ}$ are shown. The $x_{B}$ and 
$-t$-dependencies are compared to theoretical calculations performed by 
S.~Liuti and K.~Taneja \cite{simonetta_2}. The model relies on the impulse 
approximation and uses advanced spectral functions of nuclei. The calculations 
are at slightly different kinematics than our data but still provide some 
guidance. The experimental results appear to have larger asymmetries than the 
calculations. These differences may arise from nuclear effects, such as 
long-rage interactions, which are not taken into account in the model.

@@

The $Q^2$, $X_B$, and $t$ dependencies of the $^4$He CFF $\mathcal{H}_A$ 
extracted from the fit to the azimuthal dependence of $A_{LU}$ are shown in 
Figure \ref{fig:CFF_HA}. Curves on the graphs are model calculations, labelled 
{\it convolution} and {\it off-shell}. In the convolution model 
\cite{Vadim_priv}, the nucleus is assumed to be composed of non-relativistic 
and non-interacting nucleons, where these nucleons interact independently with 
the probe. The Convolution-Dual model is based on nucleon GPDs from the dual 
parametrization \cite{Guzey:2006xi}, where the Convolution-VGG uses nucleon 
GPDs  from the VGG model and is based on the double distributions ansatz 
\cite{DD_model}. The off-shell model \cite{GonzalezHernandez:2012jv} relies on 
the impulse approximation and uses advanced spectral functions of  nuclei that 
account for all configurations of the final nuclear system and the binding 
effects between the nucleons.

The results in Figure \ref{fig:CFF_HA} show that the extraction of the CFF
from the beam spin asymmetry is possible without any model dependent 
assumptions. The amplitude and the dependencies observed as a function of 
$Q^{2}$, $x_B$, and $-t$ are in agreement 
with the theoretical calculations. One can see a difference between the 
precision of the extracted imaginary and real parts, which is expected from
Eq.~\ref{eq:A_LU-coh} because $\alpha_2$ is smaller than $\alpha_0$. While the 
accuracy of our results does not allow to discriminate between the models, they 
demonstrate possibility of extraction of the CFF of spin 0 target in a model 
independent way.

%conclusion
In summary, we presented the first exclusive measurement of coherent DVCS off 
$^4$He using the CLAS spectrometer supplemented with a Radial Time Projection 
Chamber and a high pressure gaseous target. This setup allowed detection of the 
low energy $^4$He recoils in order to ensure the DVCS exclusivity.
The azimuthal dependence of the measured beam spin asymmetry ($A_{LU}$) has 
been used to extract, in model independent way, the real and the imaginary 
parts of the $^4$He CFF, $\mathcal{H}_A$. The extracted CFF is in a good 
agreement with predictions of the available models. This first fully exclusive 
experiment opens new perspectives for studying the nuclear structure with the 
GPD framework and paves a way for future measurements at JLAB using 12 GeV 
CEBAF and upgraded equipment.
